D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling

### 2025-06-17 doc/article/article.tex
Article update

\subsection{Classifier Module}
\subsubsection{Preprocessing}
TODO \\

@@

\subsection{Arrow Orientation}
As we've seen in the previous section, the need to find the exact position of an arrow's head arises.
The first algorithmic approach proposed to solve this task - called \textit{Double Clustering} - was as follows: 
\begin{enumerate}
	\item Apply a clustering algorithm to the bounding box.
	\item Suppose that the biggest cluster belongs to the arrow; the other clusters are thought to contain useless elements for this task, and can be discarded.
	\item Find the keypoints of the arrow using SIFT.
	\item Find the two keypoints inside the arrow with maximum distance. Assume that these two points are closely related to the two ends of the arrow.
	\item Run a template matching procedure inside the two clusters to find the head of the arrow.
\end{enumerate}

@@

>>>>

\subsection{Classifier Module}
\subsubsection{Preprocessing}
Due to the fact that various kinds of diagrams' images could be submitted to the system, it was decided to streamline a unified preprocessing pipeline in order to reduce the differences that our classifier model had to handle. A few problems were analyzed and the final preprocessing pipeline for the classifier module is composed of the following operators, in order:
\begin{itemize}
	\item \textbf{Gray Scaling}: In order to remove colors from images, which was deemed an useless information for classifying diagrams.
	\item \textbf{Otsu Thresholding}: To binarize the image and reducing the information to handle.
	\item \textbf{Median Filtering}: Some images in our dataset - and in the use-case - were taken on checkered notebooks. The application of a median filter, with kernel size equal to three, totally removed these.
	\item \textbf{Perspective Correction}: The diagrams' images, sometimes, didn't occupy the full photo frame. Finding four keypoints (if possible) and stretching the image helped to fully employ the frame.
	\item \textbf{Padding}: This was necessary in order to keep the same net for every image. Changes in resolution would have implied changes in the number of paramters of the MLP layers.
\end{itemize}

Figure~\ref{fig:classifier_preprocessing} shows the step-by-step process of preparing an input image for the network.

\begin{figure}[H]
	\centering
	\includegraphics[width=250pt]{classifier_preprocessing.png}
	\caption{
		The operations applied by the preprocessing module.
		\textbf{(a)}: The raw image, still colored.
		\textbf{(b)}: The image in gray scale.
		\textbf{(c)}: The binarized image, thanks to Otsu.
		\textbf{(d)}: The median filter result. Some previous thin lines are now no more.
		\textbf{(e)}: The image with a perspective correction.
		\textbf{(f)}: The image in a standard size, ready to enter the network.
	}
	\label{fig:classifier_preprocessing}
\end{figure}

@@

\subsection{Arrow Orientation}
As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix A} for a full overview of the first approach).

TODO: Current approach.

\begin{appendices}

\section{Double Clustering}
The first algorithmic approach proposed to solve this task - called \textit{Double Clustering} - was as follows: 
\begin{enumerate}
	\item Apply a clustering algorithm to the bounding box.
	\item Suppose that the biggest cluster belongs to the arrow; the other clusters are thought to contain useless elements for this task, and can be discarded.
	\item Find the keypoints of the arrow using SIFT.
	\item Find the two keypoints inside the arrow with maximum distance. Assume that these two points are closely related to the two ends of the arrow.
	\item Run a template matching procedure inside the two clusters to find the head of the arrow.
\end{enumerate}

@@

\end{appendices}
